---
title: "AMS-III.AJ Methodology Walkthrough"
author: "cdmAmsIIIaj"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AMS-III.AJ Methodology Walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r kable-helper, include = FALSE}
quiet_kable <- function(...) suppressWarnings(knitr::kable(...))
```


# Introduction

`cdmAmsIIIaj` provides tidyverse-aligned helpers that translate the Clean Development Mechanism (CDM) small-scale methodology **AMS-III.AJ Recovery and recycling of materials from solid wastes** into reproducible R code. The vignette demonstrates how to assess applicability, simulate data, evaluate each numbered equation, and aggregate monitoring periods.

# Applicability Checks

```{r applicability, message=FALSE}
library(cdmAmsIIIaj)
library(dplyr)

simulated <- simulate_ams_iiiaj_dataset(n_facilities = 2, n_periods = 3, seed = 2025)

applicability_quality <- check_applicability_material_quality_iiiaj(simulated$applicability, group_cols = "facility_id")
applicability_collection <- check_applicability_collection_network_iiiaj(simulated$applicability, group_cols = "facility_id")
applicability_monitoring <- check_applicability_monitoring_plan_iiiaj(simulated$applicability, group_cols = "facility_id")

applicability_quality %>%
  left_join(applicability_collection, by = "facility_id") %>%
  left_join(applicability_monitoring, by = "facility_id")
```

# Simulated Dataset

```{r simulation}
str(simulated, max.level = 1)
```

# Equation Walkthrough

```{r equations}
baseline <- calculate_baseline_emissions_iiiaj(
  simulated$baseline,
  group_cols = "facility_id",
  days_col = "days_in_period"
)

project <- calculate_project_emissions_iiiaj(
  simulated$project,
  group_cols = "facility_id",
  days_col = "days_in_period"
)

leakage <- calculate_leakage_emissions_iiiaj(
  simulated$leakage,
  group_cols = "facility_id"
)

list(baseline = baseline, project = project, leakage = leakage)
```

# Monitoring Period Aggregation

```{r aggregation}
aggregate_monitoring_periods_iiiaj(
  simulated$baseline,
  group_cols = c("facility_id")
)
```

# Meta-Function

```{r meta}
reductions <- estimate_emission_reductions_ams_iiiaj(
  baseline_data = simulated$baseline,
  project_data = simulated$project,
  leakage_data = simulated$leakage,
  group_cols = "facility_id",
  baseline_args = list(days_col = "days_in_period"),
  project_args = list(days_col = "days_in_period")
)
reductions
```

# Function Reference

```{r reference}
tibble::tibble(
  Function = c(
    "calculate_baseline_emissions_iiiaj",
    "calculate_project_emissions_iiiaj",
    "calculate_leakage_emissions_iiiaj",
    "calculate_emission_reductions_iiiaj",
    "aggregate_monitoring_periods_iiiaj",
    "estimate_emission_reductions_ams_iiiaj",
    "check_applicability_material_quality_iiiaj",
    "check_applicability_collection_network_iiiaj",
    "check_applicability_monitoring_plan_iiiaj",
    "simulate_ams_iiiaj_dataset"
  ),
  Signature = c(
    "(data, material_processed_col, ...)",
    "(data, electricity_consumption_col, ...)",
    "(data, collected_material_col, ...)",
    "(baseline, project, leakage, ...)",
    "(data, group_cols, sum_cols = NULL)",
    "(baseline_data, project_data, leakage_data, ...)",
    "(data, material_type_col, ...)",
    "(data, segregation_rate_col, ...)",
    "(data, throughput_monitoring_col, ...)",
    "(n_facilities = 3, n_periods = 4, seed = NULL)"
  ),
  Purpose = c(
    "Implements baseline equation (1) for avoided virgin and disposal emissions.",
    "Implements project equation (2) for energy and auxiliary material emissions.",
    "Implements leakage equation (3) covering logistics and market effects.",
    "Combines equations to deliver net emission reductions.",
    "Aggregates monitoring data across reporting periods.",
    "Runs the full AMS-III.AJ estimation workflow.",
    "Tests recycled material quality applicability conditions.",
    "Evaluates collection network applicability conditions.",
    "Verifies monitoring plan applicability conditions.",
    "Generates simulated datasets for analysis and testing."
  )
) %>%
  quiet_kable()
```

# Workflow Overview

```{r workflow}
tibble::tribble(
  ~Step, ~Description,
  "Applicability", "Use the three applicability helpers to confirm project eligibility.",
  "Simulation", "Generate or ingest monitoring datasets for baseline, project, and leakage streams.",
  "Equations", "Run the equation-level helpers to compute emissions components.",
  "Aggregation", "Aggregate monitoring data to reporting periods as required.",
  "Meta", "Call `estimate_emission_reductions_ams_iiiaj()` to produce net results."
) %>%
  quiet_kable()
```
