---
title: "AMS-I.J Methodology Walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AMS-I.J Methodology Walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This vignette demonstrates how to reproduce the AMS-I.J calculations using `cdmAmsIj`. The
methodology covers solar water heating systems that displace fossil or electric water heating.
The core steps include summing useful solar thermal output, applying baseline emission factors,
and accounting for auxiliary energy use.

# Applicability Checks

```{r applicability}
library(cdmAmsIj)
check_applicability_swh_capacity(capacity_mwth = 18)
check_applicability_solar_resource(annual_irradiation_kwhm2 = 1900)
check_applicability_backup_fraction(backup_fraction = 0.15)
```

# Simulated Dataset

```{r simulation}
set.seed(123)
example_data <- simulate_ams_ij_dataset(n_sites = 4, n_periods = 6, start_year = 2024, start_month = 7)
knitr::kable(
  head(example_data),
  format = "html",
  digits = 3
)
```

The simulation includes monitoring metadata (`monitoring_period`, `year`, `month`,
`day`, and `monitoring_label`) so that calculations can be aggregated over reporting
periods while preserving site-level identifiers.

# Equation Walkthrough

```{r calculations}
useful_output <- calculate_useful_thermal_output(example_data, group_cols = "site_id")
baseline_emissions <- calculate_baseline_emissions(
  useful_output,
  baseline_emission_factor = example_data$baseline_emission_factor[1]
)
project_emissions <- calculate_project_emissions_auxiliary(
  example_data,
  energy_col = "auxiliary_energy_mwh",
  emission_factor = example_data$auxiliary_emission_factor[1],
  group_cols = "site_id"
)
emission_reductions <- calculate_emission_reductions(baseline_emissions, project_emissions)
knitr::kable(head(emission_reductions), format = "html", digits = 3)
```

# Monitoring Period Aggregation

```{r monitoring}
period_summary <- aggregate_monitoring_periods(
  thermal_data = example_data,
  monitoring_cols = c("monitoring_label"),
  group_cols = "site_id",
  useful_energy_col = "useful_heat_mwh",
  auxiliary_energy_col = "auxiliary_energy_mwh",
  baseline_factor_col = "baseline_emission_factor",
  auxiliary_factor_col = "auxiliary_emission_factor"
)

knitr::kable(head(period_summary), format = "html", digits = 3)
```

# Meta-Function

```{r meta}
estimate_emission_reductions_ams_ij(
  thermal_data = example_data,
  baseline_emission_factor = example_data$baseline_emission_factor[1],
  auxiliary_emission_factor = example_data$auxiliary_emission_factor[1],
  group_cols = "site_id"
)
```

# Function Reference

```{r reference}
function_reference <- tibble::tibble(
  Function = c(
    "`calculate_useful_thermal_output()`",
    "`calculate_baseline_emissions()`",
    "`calculate_project_emissions_auxiliary()`",
    "`calculate_emission_reductions()`",
    "`aggregate_monitoring_periods()`",
    "`estimate_emission_reductions_ams_ij()`",
    "`simulate_ams_ij_dataset()`"
  ),
  Signature = c(
    "`calculate_useful_thermal_output(thermal_data, energy_col = \"useful_heat_mwh\", group_cols = NULL)`",
    "`calculate_baseline_emissions(useful_output, baseline_emission_factor, output_col = \"baseline_emissions_tco2e\")`",
    "`calculate_project_emissions_auxiliary(auxiliary_data, energy_col = \"auxiliary_energy_mwh\", emission_factor = 0, group_cols = NULL)`",
    "`calculate_emission_reductions(baseline_emissions, project_emissions, baseline_col = \"baseline_emissions_tco2e\", project_col = \"project_emissions_tco2e\", output_col = \"emission_reductions_tco2e\")`",
    "`aggregate_monitoring_periods(thermal_data, monitoring_cols = c(\"year\", \"month\"), group_cols = \"site_id\")`",
    "`estimate_emission_reductions_ams_ij(thermal_data, baseline_emission_factor, auxiliary_emission_factor = 0, group_cols = NULL)`",
    "`simulate_ams_ij_dataset(n_sites = 10, n_periods = 12, start_year = 2023, start_month = 1, mean_useful_heat_mwh = 360, sd_useful_heat_mwh = 55, baseline_emission_factor = 0.22, auxiliary_emission_factor = 0.18)`"
  ),
  `Equation (LaTeX)` = c(
    "\\(Q_{useful,mp} = \\sum_{i} H_{i,mp}\\)",
    "\\(E_{BL,mp}^{CO2} = Q_{useful,mp} \\times EF_{BL}\\)",
    "\\(E_{PR,mp}^{CO2} = Q_{aux,mp} \\times EF_{PR}\\)",
    "\\(ER_{mp} = E_{BL,mp}^{CO2} - E_{PR,mp}^{CO2}\\)",
    "\\(ER_{mp}^{agg} = \\sum_{i} ER_{i,mp}\\)",
    "\\(ER_{total} = \\sum_{mp} ER_{mp}\\)",
    "\\(H_{i,mp} \\sim \\mathcal{N}(\\mu_{H}/P, \\sigma_{H}/\\sqrt{P})\\)"
  ),
  Purpose = c(
    "Summed useful solar thermal output for each group or monitoring period.",
    "Baseline emissions derived from displaced conventional water heating.",
    "Project emissions reflecting auxiliary heating energy.",
    "Emission reductions obtained from the difference between baseline and project emissions.",
    "Monitoring-period aggregation consistent with methodology equations.",
    "Meta-calculation wrapper chaining the equation helpers.",
    "Data generator for examples, testing, and onboarding."
  )
)

knitr::kable(function_reference, format = "html", escape = FALSE)
```
