---
title: "AMS-I.C Methodology Walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AMS-I.C Methodology Walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This vignette demonstrates how to reproduce the AMS-I.C calculations using `cdmAmsIc`. The
methodology covers renewable thermal energy production (e.g. biomass boilers, solar thermal
systems, geothermal heat) that displaces fossil-fuel-based heat generation. The core steps
involve calculating baseline thermal output, estimating emissions using fossil heat factors,
and incorporating any residual project emissions.

# Applicability Checks

```{r applicability}
library(cdmAmsIc)
check_applicability_thermal_capacity(capacity_mwth = 12)
check_applicability_renewable_supply(renewable_fraction = 0.85)
check_applicability_fossil_displacement(fossil_heat_share = 0.75)
```

# Simulated Dataset

```{r simulation}
set.seed(123)
example_data <- simulate_ams_ic_dataset(n_facilities = 4, n_periods = 6, start_year = 2024, start_month = 7)
knitr::kable(
  head(example_data),
  format = "html",
  digits = 3
)
```

The simulation includes monitoring metadata (`monitoring_period`, `year`, `month`,
`day`, and `monitoring_label`) so that calculations can be aggregated over reporting
periods while preserving facility-level identifiers.

# Equation Walkthrough

```{r calculations}
baseline_output <- calculate_baseline_thermal_output(example_data, group_cols = "facility_id")
baseline_emissions <- calculate_baseline_emissions(
  baseline_output,
  baseline_emission_factor = example_data$baseline_emission_factor[1]
)
project_emissions <- calculate_project_emissions(
  baseline_output,
  project_emission_factor = example_data$project_emission_factor[1]
)
emission_reductions <- calculate_emission_reductions(baseline_emissions, project_emissions)
knitr::kable(head(emission_reductions), format = "html", digits = 3)
```

# Monitoring Period Aggregation

```{r monitoring}
period_summary <- aggregate_monitoring_periods(
  thermal_data = example_data,
  monitoring_cols = c("monitoring_label"),
  group_cols = "facility_id",
  energy_col = "thermal_energy_mwh",
  baseline_factor_col = "baseline_emission_factor"
)

knitr::kable(head(period_summary), format = "html", digits = 3)
```

# Meta-Function

```{r meta}
estimate_emission_reductions_ams_ic(
  thermal_data = example_data,
  baseline_emission_factor = example_data$baseline_emission_factor[1],
  project_emission_factor = example_data$project_emission_factor[1],
  group_cols = "facility_id"
)
```

# Function Reference

```{r reference}
function_reference <- tibble::tibble(
  Function = c(
    "`calculate_baseline_thermal_output()`",
    "`calculate_baseline_emissions()`",
    "`calculate_project_emissions()`",
    "`calculate_emission_reductions()`",
    "`aggregate_monitoring_periods()`",
    "`estimate_emission_reductions_ams_ic()`",
    "`simulate_ams_ic_dataset()`"
  ),
  Signature = c(
    "`calculate_baseline_thermal_output(thermal_data, energy_col = \"thermal_energy_mwh\", group_cols = NULL)`",
    "`calculate_baseline_emissions(baseline_output, baseline_emission_factor, output_col = \"baseline_emissions_tco2e\")`",
    "`calculate_project_emissions(baseline_output, project_emission_factor = 0, output_col = \"project_emissions_tco2e\")`",
    "`calculate_emission_reductions(baseline_emissions, project_emissions, baseline_col = \"baseline_emissions_tco2e\", project_col = \"project_emissions_tco2e\", output_col = \"emission_reductions_tco2e\")`",
    "`aggregate_monitoring_periods(thermal_data, monitoring_cols = c(\"year\", \"month\"), group_cols = \"facility_id\")`",
    "`estimate_emission_reductions_ams_ic(thermal_data, baseline_emission_factor, project_emission_factor = 0, group_cols = NULL)`",
    "`simulate_ams_ic_dataset(n_facilities = 10, n_periods = 12, start_year = 2023, start_month = 1, mean_thermal_mwh = 12000, sd_thermal_mwh = 1800, baseline_emission_factor = 0.27, project_emission_factor = 0.01)`"
  ),
  Purpose = c(
    "Summed baseline thermal output for each group or monitoring period.",
    "Baseline emissions using the displaced fossil heat factor.",
    "Project emissions representing auxiliary fossil energy use.",
    "Emission reductions for each monitoring period.",
    "Aggregates monitoring data to reporting periods.",
    "Meta-calculation composing the core equations.",
    "Simulation with monitoring metadata for testing and demos."
  )
)

knitr::kable(function_reference, format = "html", escape = FALSE)
```
